\LARGE \bf Artificial Cognition for Social Human-Robot Interaction:\\ An Implementation

### 2016-06-30 paper.tex
Update on Overleaf.
Work on conclusion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multi-Modal Communication}
\label{sect|com}

@@

Natural language is a basic interaction modality that we use in our system both
as an input (processing of the human speech) and as an output (verbalization of the
robot intentions). Natural language processing is facilitated by our architecture
that processes information with semantics that are already close to a
human-level. This section presents the main features of our speech processor,
{\sc Dialogs}, that include semantic and multi-modal grounding, and interactive
disambiguation. Algorithmic and implementation details are provided
in~\cite{Lemaignan2011a}.

@@

\subsubsection{Multi-Modality}

@@

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. Our architecture uses the HATP
planner (\emph{Human Aware Task Planner})~\cite{Alili2008,
Alili2009,Lallement2014}.

The HATP planning framework extends the traditional Hierarchical Task
Network (HTN) planning domain representation and semantics by making them more
suitable to produce plans which involve humans and robots acting together
toward a joint goal. HATP is used by the robot to produce human-robot
\emph{shared plans}~\cite{Grosz1996,Clark1996,Kemp2007} which are then used to
anticipate human action, to suggest a course of action to humans, or even to ask
help from the human.

The HATP planning domain defines a set of methods describing how to
decompose a task and represents the procedural knowledge of the robot as well as
its knowledge about the actions that the human partner is able to achieve. It is
stored outside of the central knowledge base, using a specific formalism (see
the related discussion at the end of this section).

@@

The originality of HATP resides in its ability to produce plans for the robot's
actions as well as for the other participants (humans or robots), that we
call \emph{shared plans}.

@@

This effectively enriches the interaction capabilities of the robot by providing
the system with what is in essence a prediction of the human behaviour. This
prediction is used by the robot execution controller to monitor the engagement
of the human partner during joint tasks.

The planner also generates synchronisation points between the agents. For
instance, in the plan depicted on Figure~\ref{plan_hatp1}, the human needs to
wait for the success of the robot's action {\tt PUTRV}. The robot monitors the
success of its own action (by checking for the fulfilment of the action
post-conditions; in this particular case \stmt{HUMAN sees BOTTLE} and
\stmt{HUMAN canReach BOTTLE}) to estimate what and when the human is likely to
perform his next action (here, \concept{TAKE(BOTTLE, TABLE)}). This, in turn,
allows the robot to monitor the human engagement and progress in the shared
plan.

We also use shared plans to verbalise the sequence of actions and to explain
the human partner how a task may be shared~\cite{warnier2012when}.

@@

\caption{An example of plan produced by HATP for a task consisting in
    cooperatively moving objects into their associated bins. Two action streams
    are generated (human actions at the top, robot actions at the bottom, {\sf
    PUTRV} stands here for {\it Put it so it is both reachable and visible}).
    The arrow between the two streams represents a synchronization point.}

@@

A cost and a duration function are associated to each action.  The duration
function provides a duration interval for the action achievement and is used, in
one hand, to schedule the different streams and, on the other hand, as an
additional cost function.

@@

We have presented in this paper a large deliberative architecture designed for
social robots. While most of its sub-components have been independently
presented in other publications, we offer here for the first time a perspective
on the overall integration of these components into a coherent and consistent
system for social human-robot interaction. We have first exposed our underlying
knowledge model based on Description Logics~\cite{Lemaignan2010} and some of the
resulting reasoning capabilities pertaining to disambiguation~\cite{Ros2010b}
and mental modelling~\cite{warnier2012when} that are shown to effectively
scaffold interaction using human-level semantics and cognitive skills. We have
then presented our approach to symbol grounding, build on an amodal situation
assessment environment~\cite{Sisbot2011} that supports perspective
taking~\cite{Marin2008,Ros2010}. We
combine it further with an advanced natural
language processor~\cite{Lemaignan2011a} to provide complete multi-modal interactive
communication.  The paper also covers our symbolic social task
planner~\cite{Alili2008, Alili2009, Lallement2014}: it generates predictive
plans of the human actions that enable the system to plan for joint human-robot
tasks. It can also make use of social heuristics to optimize plans for social
acceptability. We briefly mention our human-aware motion and manipulation
planner~\cite{Sisbot2008,Mainprice2011,Pandey2011}, and finally present two
execution controller, the PRS-based {\sc Shary}~\cite{clodic2008shary,fiore2014} and the
event-driven {\sc pyRobots}~\cite{lemaignan2015pyrobots}.

The integration of these components in a consistent, working and observable
system is supported by the design of our software interfaces: most of the
datastreams exchanged by our components uses high-level semantics in the form of
first-order logic statements. The two reported support studies further evidence
how our use of explicit, high-level semantics provide a low-impedance,
human-level mean of building a complex deliberative
architecture for socially interactive robots~\cite{Lemaignan2013}.

\fixme{Rachid -- TDB: what are the other 'deliberative architecture for socially interactive
robots' and where/how ours is better?}

While several contributions in the literature provide insights and contributions 
on one aspect or another, we are not aware of a fully implemented architecture that effectively
combines in a coherent manner all the points mentioned above.
We do not claim that this represents the only architecture that can be designed
to address such problems, nor that our approach fully addresses each and every
of these issues. Our point here is that it is a fully implemented architecture that allowed us to
identify, refine and combine a number of key cognitive abilities that we believe are necessary. Indeed, 
we show here below that it provided us a mean to develop and explore a number of key features for HRI 
in terms of models and algorithms and their articulation in an integrated decisional system for a 
cognitive and interactive robot.

\fixme{Discussion on deliberative architectures != cognitive architecture}: a \emph{cognitive
architecture} is usually understood as an artificial yet principled model of (human)
cognition, while we propose a practical approach to consistently combine a large
set of independent cognitive processes by bridging them through high-level semantics.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multi-Modal Communication and Situated Dialogue}
\label{sect|com}

@@

Natural language is a basic interaction modality that we use in our system both
as an input (processing of the human speech) and as an output (verbalization of the
robot intentions as well human-robot shared plans). Natural language processing is facilitated by our architecture
that processes information with semantics that are already close to a
human-level. This section presents the main features of our speech processor,
{\sc Dialogs}, that include semantic and multi-modal grounding, and interactive
disambiguation. Algorithmic and implementation details are provided
in~\cite{Lemaignan2011a}.

@@

\subsubsection{Dealing with Multi-Modality}

@@

Whenever necessary, the execution controllers rely on symbolic task planning to convert long-term
desires into a set of partially ordered elementary actions. This is the role of the HATP
planner (\emph{Human Aware Task Planner})~\cite{Alili2008,
Alili2009,Lallement2014}.

The HATP planning framework extends the traditional Hierarchical Task
Network (HTN) planning domain representation and semantics by making them more
suitable to produce plans which involve humans and robots acting together
toward a joint goal. HATP is used by the robot to produce human-robot
\emph{shared plans}~\cite{Grosz1996,Clark1996,Kemp2007} which are then used to
anticipate human action, to suggest a course of action to humans, or even to ask
help from the human if needed.

The HATP planning domain defines a set of methods describing how to
incrementally decompose a task and to allocate subtasks and actions to the robot and/or the human depending on the context. This represents the procedural knowledge of the robot as well as
its knowledge about the actions that the human partner is able to achieve. It is
stored outside of the central knowledge base, using a specific formalism (see
the related discussion at the end of this section).

We briefly discuss here below how HATP incrementally builds and synchronises streams of actions concerning each or the agents (humans and robot) involved in a task, and how it promotes plans that satisfy humans needs and preferences as well as comfort and legibility issues.

@@

The originality of HATP resides in its ability to produce \emph{shared plans} which might involve 
the robot as well as the other participants (humans or robots).

@@

This effectively enriches the interaction capabilities of the robot by providing
the system with what is in essence a prediction of the human behaviour. This
prediction is used by the robot execution controller to monitor the engagement
of the human partner during plan achievement.

The planner also generates causal links and synchronisation points between the robot and his human partners. For
instance, in the plan depicted on Figure~\ref{plan_hatp1}, the human needs to
wait for the success of the robot's action {\tt PUTRV}. The robot monitors the
success of its own action (by checking for the fulfilment of the action
post-conditions; in this particular case \stmt{HUMAN sees BOTTLE} and
\stmt{HUMAN canReach BOTTLE}) to estimate what and when the human is likely to
perform his next action (here, \concept{TAKE(BOTTLE, TABLE)}). This, in turn,
allows the robot to monitor the human engagement and the progress in the shared
plan execution.

We also use shared plans to verbalise the sequence of actions and to explain
the human partner how a task may be shared~\cite{warnier2012when,fiore2016planning,milliez2016using}.

@@

\caption{An example of plan produced by HATP for a task consisting in
    cooperatively moving objects into their associated bins. Two action streams
    are generated (human actions at the top, robot actions at the bottom, {\sf
    PUTRV} stands here for {\it Put it so it is both Reachable and Visible}).
    The arrow between the two streams represents a synchronization point between 
    the robot and the human based on a causal link.}

@@

A cost and a duration function are associated to each action.  The duration
function provides a duration interval for the action achievement and is used, in
one hand, to schedule the different streams and, on the other hand, as an
additional cost function. Cost here is not used here to estimate only standard time and energy consumption but
to estimate a number factors that measure the satisfaction of the human in terms of acceptability and legibility of robot behavior.

@@

We have presented in this paper a large deliberative architecture designed for
social robots. While most of its sub-components have been independently
presented in other publications, we offer here for the first time a perspective
on the model of integration of these components into a coherent and consistent
system for social human-robot interaction. We have first exposed our underlying
knowledge model based on Description Logics~\cite{Lemaignan2010} and some of the
resulting reasoning capabilities pertaining to disambiguation~\cite{Ros2010b}
and mental modelling~\cite{warnier2012when} that are shown to effectively
scaffold interaction using human-level semantics and cognitive skills. We have
then presented our approach to symbol grounding, build on an amodal situation
assessment environment~\cite{Sisbot2011} that supports perspective
taking~\cite{Marin2008,Ros2010}. We combine it further with a situated natural
language processor~\cite{Lemaignan2011a} to provide complete multi-modal
interactive communication.  The paper also covers our symbolic social task
planner~\cite{Alili2008, Alili2009, Lallement2014}: it generates predictive
plans of the human actions that enable the system to plan for joint human-robot
tasks. It can also make use of social heuristics to optimize plans for social
acceptability. We briefly mention our human-aware motion and manipulation
planner~\cite{Sisbot2008,Mainprice2011,Pandey2011}, and finally present two
execution controller, the PRS-based {\sc Shary}~\cite{clodic2008shary,fiore2014}
and the event-driven {\sc pyRobots}~\cite{lemaignan2015pyrobots}.

The integration of these components in a consistent, working and observable
system builds upon the particular design of the interfaces between the
cognitive components: the information streams use high-level semantics,
represented as first-order logic statements. In that sense, our deliberative
architecture is similar to projects like {\sc cram}~\cite{Beetz2010} or PEIS
Ecology~\cite{saffiotti2005peis, daoutis2012cooperative}, with however a
stronger emphasize on the representation and interaction with humans. We
however distinguish ourselves from research on \emph{cognitive architectures}:
cognitive architectures are usually understood as an artificial yet principled
model of (human) cognition. While some have been deployed on autonomous robots,
like HAMMER~\cite{demiris2006hierarchical} or ACT-R/E~\cite{trafton2013act},
most are primarily concerned with the modeling of human cognition and are less
focused on practical deployment on socially interactive robots.

In that sense, our contribution in terms of architecture is a practical one:
our integration model enables to consistently combine a large set of
technically independent yet cognitively interdependent cognitive processes. We
bridge them through explicit, human-level semantics, and we show that this
results in a fully implemented system, effectively deployed on several
platforms and in several real interaction scenarios.

Besides, this work introduces several new contributions related to the
representation and the management of humans in an autonomous robotic system.
Specifically, we mentioned in the introduction the following four points:

	\item our system achieves multi-modal and interactive grounding in complex
		real environments involving one or several humans and a robot;

	\item it supports a distributed computation of symbolic knowledge for
		situated dialogue, thanks to the combination of perspective taking,
		affordances computation and logical inference;

	\item it provides generic mechanisms for the robot to reason about the mental state
		of its human partners;

	\item and, by reusing the computed affordances and inference, it generates,
		monitors and takes part to human-robot shared plans;

While several contributions in the literature provide insights and
contributions on one aspect or another (references are in the corresponding
subsections), we are not aware of a fully implemented architecture that
effectively combines in a coherent manner all these points. Our approach
allowed us to identify, refine and combine a number of key cognitive abilities
that we believe are needed for successful and autonomous social human-robot
interaction. The novelty and relevance of this contribution to HRI is
underlined by the range of multi-disciplinary collaborations and studies that
have been made possible by our
architecture~\cite{dautenhahn2006may,koay2007exploratory,Ros2010b,dehais2011physiological,ferreira2015users,GharbiROMAN2015}.
